Modal Logics in the Coq Proof Assistant

### 2014-12-10 Papers/InProgress/ModalLogicCoq/ModalCoq.tex
Write section about modal logics beyond K

\section{Modal Logics beyond K}

\noindent
The most well-known modal logics can be distinguished by their acceptance of the following reachability axioms. \textbf{S5} is the modal logic that accepts all three.

\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
\Axiom reflexivity: \fforall w, r w w.
\Axiom transitivity: \fforall w1 w2 w3, (r w1 w2) -> (r w2 w3) -> (r w1 w3).
\Axiom symmetry: \fforall w1 w2, (r w1 w2) -> (r w2 w1).
\end{Verbatim}

\noindent
Modal logic ``axioms'' can be derived from the reachability axioms.

\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
\Theorem K: 
  [ mforall p, mforall q, (box (p m-> q)) m-> (box p) m-> (box q) ].
\Proof. mv. 
intros p q H1 H2. box_i. box_e H1 H3. apply H3. box_e H2 H4. exact H4. 
\Qed.

\Theorem T: [ mforall p, (box p) m-> p ].
\Proof. mv. 
intro p. intro H. box_e H H1. exact H1. apply reflexivity. 
\Qed.
\end{Verbatim}

\noindent
In strong modal logics, such as S5, iterations of modal operators can be collapsed. This is a controversial principle used in G\"odel's and Scott's versions of the proof.

>>>>

\section{Modal Logics beyond K}
\label{sec:BeyondK}

The embedding described in Section \ref{sec:Embedding} and the 
new tactics described in Section \ref{sec:Tactics} allow convenient 
interactive reasoning for modal logic \textbf{K} within \Coq. 
The axiom K is easily derivable:

\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
\Theorem K: 
  [ mforall p, mforall q, (box (p m-> q)) m-> (box p) m-> (box q) ].
\Proof. mv. 
intros p q H1 H2. box_i. box_e H1 H3. apply H3. box_e H2 H4. exact H4. 
\Qed.
\end{Verbatim}

\noindent
For other modal logics beyond \textbf{K}, their frame conditions constraining 
the reachability relation must be stated as \Coq axioms.

\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
\Axiom reflexivity: \fforall w, r w w.

\Axiom transitivity: \fforall w1 w2 w3, (r w1 w2) -> (r w2 w3) -> (r w1 w3).

\Axiom symmetry: \fforall w1 w2, (r w1 w2) -> (r w2 w1).
\end{Verbatim}

\noindent
Hilbert-style modal logic axioms, such as for example T, 
can be easily derived from their corresponding frame conditions:

\begin{Verbatim}[commandchars=\\\{\},fontsize=\verbsize]
\Theorem T: [ mforall p, (box p) m-> p ].
\Proof. mv. 
intro p. intro H. box_e H H1. exact H1. apply reflexivity. 
\Qed.
\end{Verbatim}

\noindent
In strong modal logics, such as S5 (which enforces all three frame conditions specified above), iterations of modal operators can be collapsed. This is a controversial principle used in G\"odel's proof shown in Section \ref{sec:proof}.
